---
title: "Weather vs. Climate"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    
---

```{r setup, include=FALSE}


# Title: Weather vs. Climate
# Data Source: NOAA - Global Historical Climatology Newtork (GHCN)
# Date: Started in Summer 2019, First Shared on Github 11 Feb. 2020
# Created By: Ryan McKeon - Geography Department | Dartmouth College | ryan.mckeon@gmail.com




# Import Libraries
library(flexdashboard)
library(rnoaa)
library(data.table)
library(lubridate)
library(sf)
library(spData)
library(igraph)
library(leaflet)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(zoo)
library(knitr)
library(gridExtra)

# Set some Parameters:

# Pick A Weather Station  # MAKE DYNAMIC IN SHINY
stationID = "USC00273850"  # Hanover NH
#stationID = "USC00471416"  # Hanover NH

# Pick a year for temperature plot
currentYear = 2016  # MAKE DYNAMIC IN SHINY

# Path to GHCN Weather Stations Text File
GHCN_all = fread("GHCN_Weather_Stations_ALL.txt") 

# ***********************************  Filter Stations and Plot **************************

# filter all to unique station IDs for all GHCN weather stations
GHCN_stations = unique(GHCN_all, by = "Station")

# filter for active stations only
GHCN_stations = GHCN_stations[GHCN_stations$Tend > "2018"]

# filter for 30 year records of data collection
# Add a new field (Diff_yrs) and set the value equal to the difference in Tend and Tstart
GHCN_stations[, Diff_yrs := GHCN_stations$Tend - GHCN_stations$Tstart]
# subset
GHCN_select = GHCN_stations[GHCN_stations$Diff_yrs >29]

# Make an sf object out of the unique stations 
GHCN_select_shp <- st_as_sf(GHCN_select, coords = c("Long", "Lat"), crs = 4326)

# clip the data set down to the US
# get US geometry from spdata "world" data set
USonly = world[world$name_long == "United States", ]

# clip the dust GHCN stations data to only those stations in the US
USstations_shp = GHCN_select_shp[USonly, ]

# Find the station type by looking at the first three letters of the Station ID, then subset to keep only USC stations
USstations_shp$StationType = substr(USstations_shp$Station, 1,3)
USCs_shp = USstations_shp[USstations_shp$StationType == "USC", ]

# Pick a Station, Get Data, Process Data  ***********************************



# make and sf object out of the selected station
current_station_shp <- USCs_shp[USCs_shp$Station == stationID, ]


# Beacuse not all variables have the same number of data entries... 

# create an unbroken range of dates from the first observations to modern times and make it into the first row of a data frame
#date = seq.Date(as.Date("1893/6/1"), as.Date("2019/6/1"), "day")
#WXdf = data.frame(date)

# Get the daily data for weather station in separate data frames
TempMin = as.data.frame(ghcnd_search(stationID, var = "TMIN", refresh = TRUE))  # units = 1/10 of degC   , refresh = TRUE
TempMAX = as.data.frame(ghcnd_search(stationID, var = "TMAX", refresh = TRUE))  # units = 1/10 of degC
PrecipMM = as.data.frame(ghcnd_search(stationID, var = "PRCP", refresh = TRUE)) # units = 1/10 of mm
#SnowFall = as.data.frame(ghcnd_search(stationID, var = "SNOW")) # mm
#SnowDepth = as.data.frame(ghcnd_search(stationID, var = "SNWD")) # mm
#dustObs = as.data.frame(ghcnd_search(stationID, var = "WT07")) # dust obs


# Match dates to organize data and convert to deg C and mm

# %%%%%%%%%%%%%%%%% Hotness and Coldness Indicies %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# from https://eos.org/geofizz/hotness-and-coldness-indexes-based-on-the-fahrenheit-scale

# H-index for Max temp
TempMAX$degF = round(TempMAX$tmax.tmax/10*(9/5)+32)  # create a new deg F column and round to nearest degree
TempMAX$year = year(TempMAX$tmax.date)  # create a year column for looping through
startYear = min(TempMAX$year)
endYear = max(TempMAX$year)
lengthOfRecord = endYear - startYear
hFactor = setNames(data.frame(matrix(ncol = 5, nrow = lengthOfRecord)), c("year","numDays","nNAs","hNum", "cNum"))

# C index for Min Temp
TempMin$degF = round(TempMin$tmin.tmin/10*(9/5)+32)  # create a new deg F column and round to nearest degree
TempMin$coldness = TempMin$degF-32
TempMin$year = year(TempMin$tmin.date)


rowNum = 0

for (i in startYear:endYear) {
  rowNum = rowNum + 1
  hFactor[rowNum,"year"] = i
  yearMax = TempMAX[TempMAX$year == i, ]  # get one year of data
  
  # Find the number of days in the year and scan for NAs 
  hFactor[rowNum,"numDays"] = nrow(yearMax)  # number of days in the record for the year
  gotNAs = is.na(yearMax$degF)   # find the number of NAs for that year
  hFactor[rowNum,"nNAs"] = length(gotNAs[gotNAs == TRUE])
  
  # H Index calculation
  yearMax$rank = rank(-yearMax$degF, ties.method = "random")  # rank the high temps in decsending order
  yearMax$test <- as.numeric(yearMax$rank < yearMax$degF)  # create new column to store comparsion of rank with temp 
  yearMax = yearMax[yearMax$test == 1, ]  # subset to keep only those rows where rank > temp
  hFactor[rowNum,"hNum"]  = max(yearMax$rank, na.rm = TRUE) # find the crossover where rank exceeds temperature --  the H factor
  
  # C Index Calculation
  yearMin = TempMin[TempMin$year == i, ]
  yearMin$rank = rank(yearMin$coldness, ties.method = "random")
  yearMin$test <- as.numeric(yearMin$rank < -yearMin$coldness)
  yearMin = yearMin[yearMin$test == 1, ]
  hFactor[rowNum,"cNum"]  = max(yearMin$rank, na.rm = TRUE)
}

# Screen for incomplete years or those with too many missing data points
hFactor <- hFactor %>%
  filter(numDays > 365-(365*0.05))

hFactor <- hFactor %>%
  filter(nNAs < 365*0.05)    # remove years with greater than 5% of the data missing (i.e. NAs)
     # remove years where more than 5% of the days are missing from the record

xLow = min(hFactor$year)
xHigh = max(hFactor$year)

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% End Hotness and Coldness Indicies %%%%%%%%%%%%%%%%%%%%%%


# *********************  Create Climographs for 30 yr normals ****************************************

# combine temperature data tables with dplyr left_join(), note that it allows you to match join fields of different names
TempALL = left_join(TempMAX, TempMin, by = c("tmax.date" = "tmin.date"))  # note that order matters... Tmax is ".x", Tmin is ".y" for shared column names
TempPreALL = left_join(TempALL, PrecipMM, by = c("tmax.date" = "prcp.date"))
TempPreALL$precipInches = TempPreALL$prcp.prcp/254

# add Month, julian day, and average daily temp columns
TempPreALL$month = month(TempPreALL$tmax.date)
TempPreALL$julDay = yday(as.Date(TempPreALL$tmax.date))
TempPreALL$dailyMeanTemp = (TempPreALL$degF.x + TempPreALL$degF.y)/2


# monthly means for the whole record ==========================================
# Max Temp
outTmaxMonth = TempPreALL %>% group_by(day=floor_date(tmax.date, "month")) %>%
  summarize(degF.x=mean(degF.x))

# Min Temp
outTminMonth = TempPreALL %>% group_by(day=floor_date(tmax.date, "month")) %>%
  summarize(degF.y=mean(degF.y))

# Precip
outPrecipMonth = TempPreALL %>% group_by(day=floor_date(tmax.date, "month")) %>%
  summarize(precipInches=sum(precipInches))

# end monthly means for the whole record ======================================

# get 30 years of data for the common "Normal" 1981 - 2010 -- MAKE RANGE DYNAMIC IN SHINY
normStartYear = 1981
normEndYear = 2010
normLengthYears = normEndYear - normStartYear
norm30 = TempPreALL[TempPreALL$year.x >= normStartYear & TempPreALL$year.x <= normEndYear, ]

# create data frames for output from loops below
month_summary = setNames(data.frame(matrix(ncol = 4, nrow = 12)), c("month", "Tmin", "Tmax", "Precip"))
day_summary = setNames(data.frame(matrix(ncol = 13, nrow = 366)), c("day", "Tmin", "Tmax", "Tmean","Tmin2SD","Tmax2SD","meanPrecip","cumSumPrecip","precipSD","TmaxRecord","TminRecord","isRecordMax","isRecordMin"))


# Loop Through Months and calculate average high and low temps and sum of precip

for (i in 1:12) {
  monthDF = norm30[norm30$month == i, ] # subset the data to a single month
  month_summary[i,1] = i
  month_summary[i,2] = mean(monthDF$degF.y, na.rm = TRUE) # Tmin mean
  month_summary[i,3] = mean(monthDF$degF.x, na.rm = TRUE) # Tmax mean
  month_summary[i,4] = sum(monthDF$precipInches, na.rm = TRUE)/normLengthYears # precip sum (averaged over the length of the record)
  month_summary[i,5] = sum(monthDF$precipInches, na.rm = TRUE)
}

# Loop Through Julian Days of the year and calculate a lot of things
for (i in 1:366) {
  dayDF = TempPreALL[TempPreALL$julDay == i, ]  # subset the data to a single julian day
  day_summary[i,1] = i
  day_summary[i,2] = mean(dayDF$degF.y, na.rm = TRUE)  # mean low temp
  day_summary[i,3] = mean(dayDF$degF.x, na.rm = TRUE)  # mean high temp
  day_summary[i,4] = (day_summary[i,2] + day_summary[i,3])/2   # average the high and low to generate a crude mean for coloring lines on the plot
  day_summary[i,5] = mean(dayDF$degF.y, na.rm = TRUE) - (2*sd(dayDF$degF.y, na.rm = TRUE)) # two sigma uncertainty of the mean low temp
  day_summary[i,6] = mean(dayDF$degF.x, na.rm = TRUE) + (2*sd(dayDF$degF.x, na.rm = TRUE)) # ditto for high
  day_summary[i,7] = mean(dayDF$precipInches, na.rm = TRUE)  # mean precip
  day_summary[i,9] = sd(dayDF$precipInches, na.rm = TRUE)
  day_summary[i,10]= max(dayDF$degF.x, na.rm = TRUE)   # get the daily record high temp
  day_summary[i,11]= min(dayDF$degF.y, na.rm = TRUE)   # ditto for low temp
}

day_summary$cumSumPrecip = cumsum(day_summary$meanPrecip)  # calculate the cumulative sum of the average rainfall for each day


# get a specific year of data 

currentYearDF = TempPreALL[TempPreALL$year.x == currentYear, ]
currentYearDF$Tmean = (currentYearDF$degF.x + currentYearDF$degF.y)/2
currentYearDF$precipInches[is.na(currentYearDF$precipInches)] <- 0   # replace NAs in precip data with zeros
currentYearDF$cumSumPrecip = cumsum(currentYearDF$precipInches)



```


Row
-----------------------------------------------------------------------

### **Map of Weather Stations from the Global Historical Climatology Network** - The blue pin indicates the location of the current station being plotted below. The darkness of the point indicates the length of the record for each station. The map is dynamic, pan, zoom and click on points to get the station ID number.

```{r}
# Plot Stations with Leaflet
coords = st_coordinates(current_station_shp)
currentLat = coords[2]
currentLon = coords[1]

# color ramp for sumGrid -- Also Applied to peak prominence!
pal <- colorNumeric(
  palette = "Blues",
  domain = USCs_shp$Diff_yrs
)

Lmap <- leaflet(USCs_shp, height = 800) %>%  # create the map and set the height of the window for it
  setView(lng = currentLon, lat = currentLat, zoom = 6) %>%   # center the map on the current station
  addProviderTiles(providers$CartoDB.Positron) %>%  # add CardoDB Map Tiles
  addCircleMarkers(data = USCs_shp,
    radius=5, 
    fill = TRUE, 
    stroke = TRUE, 
    color = "black", 
    weight = 1, 
    fillColor = ~pal(Diff_yrs), 
    fillOpacity = 1, 
    popup = ~paste("Station ID: ", Station, "<br>Record Length: ", Diff_yrs, " years")
      ) %>%
  addLegend(pal = pal, values = ~USCs_shp$Diff_yrs, title = "Data Record in Years") %>%
  addMarkers(lng=currentLon, lat=currentLat, popup = "Current Weather Station")
Lmap
```

Row
-----------------------------------------------------------------------

### **Monthly Temperture and Precipition** - This is the classic "Climograph" where average monthly temperature (left Y axis) and precipitation (right Y axis) are displayed together.  This plot averages data over a 30 year period starting in 1981 and ending in 2010 (commonly referred to as the Normal by climate scientists).
  

```{r, fig.cap="Average monthly temperature (gray bars) and Precipitation (blue line). "}


# Classic Climograph Plot %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

# create field with month names and impose an order for their display using factor()
month_summary$monthNames = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec")
month_summary$monthNames = factor(month_summary$monthNames, levels = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec"))

#x = factor(month_summary$monthNames, levels = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec"))

# get the scaling factor for the secondary y axis
yScaleFactor = max((month_summary$Tmax+month_summary$Tmin)/3)/max(month_summary$Precip)

# This makes a combo bar and line plot with a second axis...  
Climograph = ggplot(month_summary, aes(x = monthNames, group = 1)) +                           # note the "group = 1" - this tells it that there is only one value per monthName group
  geom_linerange(mapping = aes(ymin = 0, ymax = (Tmax+Tmin)/2), color = "gray", size = 10) +
  geom_line(mapping = aes(y= Precip*yScaleFactor), color = "blue", size = 3) +
  scale_y_continuous(sec.axis = sec_axis(~./yScaleFactor, name = "Precipitation (inches)")) +  # Note the reversal of the scaling factor here!
  ylab("Average Monthly Temperature (deg F)") +
  xlab("") +
  theme_light()

Climograph

```

### **Data Completeness Matrix** - This tile plot show the data density for the full length of the record for the current weather station.  Every day for the record is represented, where color ranges from blue to red it is showing the daily mean temperature, where gray or white are present means that no temperature data was recorded.

```{r, fig.cap="Blue to Red colors reflect recorded daily mean temperature, gray and white represent missing data."}

# Heat map for visualizing data completeness %%%%%%%%%%%%%%%%%%%%

completeness = ggplot(TempPreALL, aes(julDay, year.x)) +
  geom_tile(aes(fill = dailyMeanTemp), color = "white") +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = "Data Record Completeness", x = "Julian Day", y = "Year") +
  scale_y_reverse() +  # flip the y axis values
  theme_light()

completeness

```




Row
-----------------------------------------------------------------------

### **Daily Temperature Range** - This plot compares the daily temperature range for a single year (the colorful vertical lines) to the average daily high and low temperature (light gray) and the record high and low temperature observed at the current station (darker gray). Red and blue diamonds indicate when the high or low temperature record was set during the particular year being observed.    
```{r,  fig.width=9, fig.height=5, fig.cap="This shows daily temperature ranges over the course of a year vs. the average high and low (light gray) and the record high and low temperature ever recorded at this station (dark gray)."}

# Find out if any days set records for the temperature history 
for (i in 1:366) {
  if (is.na(currentYearDF[i, "degF.x"]) == FALSE) {
    if (currentYearDF[i,"degF.x"] == day_summary[i,"TmaxRecord"]) {
      day_summary[i,"isRecordMax"] = currentYearDF[i,"degF.x"]
    }
  }
  if (is.na(currentYearDF[i, "degF.y"]) == FALSE) {
    if (currentYearDF[i,"degF.y"] == day_summary[i,"TminRecord"]) {
      day_summary[i,"isRecordMin"] = currentYearDF[i,"degF.y"]
    }
  }
}

plotAnnotation = paste("Daily Temperature for",as.character(currentYear), sep = " ")

# construct daily temperature range plot with 30 yr daily normals and current year's data plotted on top

ggplot() +
  #geom_linerange(data = day_summary, mapping = aes(x = day, ymin = Tmin2SD, ymax = Tmax2SD), color = "#949494", size = 1.5) +  # 30 yr.2 sigma range
  geom_linerange(data = day_summary, mapping = aes(x = day, ymin = TminRecord, ymax = TmaxRecord), color = "#949494", size = 1) +  # daily record high and low temperature range 
  geom_linerange(data = day_summary, mapping = aes(x = day, ymin = Tmin, ymax = Tmax), color = "#b0b0b0", size = 1) +        # 30 yr mean hi/lo
  geom_linerange(data = currentYearDF, mapping = aes(x = julDay, ymin = degF.y, ymax = degF.x, color = Tmean), alpha = 0.6, size = 1) +   # observed data for particular year
  geom_point(data = day_summary, mapping = aes(x=day, y=isRecordMax), shape=23, color="black", fill="red",size=3) +
  geom_point(data = day_summary, mapping = aes(x=day, y=isRecordMin), shape=23, color="black", fill="blue",size=3) +
  scale_colour_gradient(low = "blue", high = "red", guide = "colorbar") +
  #scale_fill_identity(name = "30 year Normal High and Low Temp", guide = "legend", labels = c("95% Confidence Range", "Average Range")) +  # trying to label grays on the legend... not working
  ylab('Temperature (deg F)')+
  xlab('Julian Day')+
  ylim(-40, 120) +
  #ggtitle('Daily Temperature Range')+
  labs(color="Daily Mean Temp (F)") + # set legend title here
  annotate(geom="text", x=50, y=110, label=plotAnnotation,color="Dark Blue") +
  theme_light() +
  theme(
    legend.position = c(0.53, 0.15), 
    legend.direction = "horizontal",
    axis.title=element_text(size=14),
    )

```



Row
-----------------------------------------------------------------------

### **Hotness Index - Are the hot days getting hotter and more plentiful?**  For the Hotness index, a value of 75 for a particular year means that the maximum daily temperature equalled or exceeded 75 degrees Fahrenheit 75 times during the year. *If the Hotness index is increasing, both the number of hot days and the temperature on those days is increasing.*


```{r, fig.cap="The points show the annual variation in the Hotness Index, the red line smooths the data using a loess filter and the gray shading is the 95% CI."}
hFactorPlot = ggplot() +
  geom_point(data = hFactor, mapping = aes(x=year, y=hNum),color = "red",, shape = 21, size = 2.5) +
  #geom_line(data = hFactor, mapping = aes(x=year, y=rollmean(hNum, 25, na.pad = TRUE)), color = "red", size = 4,) +
  geom_smooth(data = hFactor, mapping = aes(x=year, y=hNum),method = stats::loess, color = "red", size = 2, alpha = 0.5) +
  xlim(xLow,xHigh) +
  xlab("Year") +
  ylab("Hotness Index") +
  theme_light() 
hFactorPlot

# cFactorPlot = ggplot() +
#   geom_point(data = hFactor, mapping = aes(x=year, y=cNum),color = "blue", fill = "blue", shape = 21, size = 3) +
#   xlim(xLow,xHigh)+ 
#   xlab("Year") +
#   ylab("Coldness Index") +
#   theme_light()

#FactorPlot = ggarrange(hFactorPlot, cFactorPlot, 
#          labels = c("Hotness Index", "Coldness Index"),
#          ncol = 2, nrow = 1)
          
#FactorPlot

```

### **Coldness Index - Are the cold days getting colder and greater in number?** The Coldness index is similar; here a value of 25 means that the minimum daily temperature equalled or exceeded 25 degrees Fahrenheit *below* the freezing point of water (at 32 deg F)  25 times during the year. *If the Coldness index is dropping year to year, it means that cold days are warmer and less frequent.* 


```{r, fig.cap="The Coldness Index is defined as the number of days C at which the Fahrenheit temperature relative to the freezing point of water (32 degF) is −C or less. The points show the annual variation, the blue line smooths the data using a loess filter and the gray shading is the 95% CI."}
cFactorPlot = ggplot() +
  geom_point(data = hFactor, mapping = aes(x=year, y=cNum),color = "blue", shape = 21, size = 2.5) +
  geom_smooth(data = hFactor, mapping = aes(x=year, y=cNum),method = stats::loess, color = "blue", size = 2, alpha = 0.5) +
  xlim(xLow,xHigh)+
  xlab("Year") +
  ylab("Coldness Index") +
  theme_light()
cFactorPlot

```
Row
-----------------------------------------------------------------------

### **Monthly Mean Temperature Anomaly** - This plot illustrates how the monthly mean temperature from the last 10 years compares to the monthly mean temperature calculated using the full record from the weather station.  The "anomaly" is calculated as the difference between the last 10 years and the full record for each month.
```{r}

# anomaly plots comparing past with present at a monthly scale

# anomaly for each month is a 10 year average that is subtracted from the monthly average for the whole data set...  the hooks are here to add more time ranges to test AND to use the Normal (1981-2010) time range.

# subset 10 year chunks from the whole data set
#old10DF = TempPreALL[TempPreALL$year.x > 1899 & TempPreALL$year.x < 1910, ]
#mid10DF = TempPreALL[TempPreALL$year.x > 1959 & TempPreALL$year.x < 1970, ]
latest10DF = TempPreALL[TempPreALL$year.x > endYear-11 & TempPreALL$year.x < endYear, ]


# create output DFs
#norm30MonthDF = setNames(data.frame(matrix(ncol = 4, nrow = 12)), c("monthNum","monthName", "TempMean","PrecipMean"))
ALLMonthDF = setNames(data.frame(matrix(ncol = 4, nrow = 12)), c("monthNum","monthName", "TempMean","PrecipMean"))
#oldMonthDF = setNames(data.frame(matrix(ncol = 6, nrow = 12)), c("monthNum","monthName", "TempMean","PrecipMean","TempAnom","PrecipAnom"))
#midMonthDF = setNames(data.frame(matrix(ncol = 6, nrow = 12)), c("monthNum","monthName", "TempMean","PrecipMean","TempAnom","PrecipAnom"))
latestMonthDF = setNames(data.frame(matrix(ncol = 6, nrow = 12)), c("monthNum","monthName", "TempMean","PrecipMean","TempAnom","PrecipAnom"))

# loop through each month and calculate monthly mean temp and total precip
for (i in 1:12) {
  # The whole data set
  currentMonthDF = TempPreALL[TempPreALL$month == i, ]
  ALLMonthDF[i,"monthNum"] = i
  ALLMonthDF[i, "TempMean"] = mean(currentMonthDF$dailyMean, na.rm= TRUE)
  ALLMonthDF[i,"PrecipMean"] = sum(currentMonthDF$precipInches, na.rm = TRUE)/10  # divide by the length (in years) to get the average monthly precip.
  
  # # Norm30 DF
  # currentMonthDF = norm30[norm30$month == i, ]
  # norm30MonthDF[i,"monthNum"] = i
  # norm30MonthDF[i,"TempMean"] = mean(currentMonthDF$dailyMean, na.rm= TRUE)
  # norm30MonthDF[i,"PrecipMean"] = sum(currentMonthDF$precipInches, na.rm = TRUE)/10  
  # 
  # # Old DF
  # currentMonthDF = old10DF[old10DF$month == i, ]
  # oldMonthDF[i,"monthNum"] = i
  # oldMonthDF[i,"TempMean"] = mean(currentMonthDF$dailyMean, na.rm= TRUE)
  # oldMonthDF[i,"PrecipMean"] = sum(currentMonthDF$precipInches, na.rm = TRUE)/10  
  # oldMonthDF[i,"TempAnom"] = oldMonthDF[i,"TempMean"] - norm30MonthDF[i,"TempMean"]
  # oldMonthDF[i,"PrecipAnom"] = oldMonthDF[i,"PrecipMean"] - norm30MonthDF[i,"PrecipMean"]
  # 
  # # Mid DF
  # currentMonthDF = mid10DF[mid10DF$month == i, ]
  # midMonthDF[i,"monthNum"] = i
  # midMonthDF[i,"TempMean"] = mean(currentMonthDF$dailyMean, na.rm= TRUE)
  # midMonthDF[i,"PrecipMean"] = sum(currentMonthDF$precipInches, na.rm = TRUE)/10
  # midMonthDF[i,"TempAnom"] = midMonthDF[i,"TempMean"] - norm30MonthDF[i,"TempMean"]
  # midMonthDF[i,"PrecipAnom"] = midMonthDF[i,"PrecipMean"] - norm30MonthDF[i,"PrecipMean"]
  
  # latest DF -- vs WHOLE DATA SET 
  currentMonthDF = latest10DF[latest10DF$month == i, ]
  latestMonthDF[i,"monthNum"] = i
  latestMonthDF[i,"TempMean"] = mean(currentMonthDF$dailyMean, na.rm= TRUE)
  latestMonthDF[i,"PrecipMean"] = sum(currentMonthDF$precipInches, na.rm = TRUE)/10  
  latestMonthDF[i,"TempAnom"] = latestMonthDF[i,"TempMean"] - ALLMonthDF[i,"TempMean"]
  latestMonthDF[i,"PrecipAnom"] = latestMonthDF[i,"PrecipMean"] - ALLMonthDF[i,"PrecipMean"]
  
}

# # create field with month names and impose an order for their display using factor()
# oldMonthDF$monthName = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec")
# oldMonthDF$monthName = factor(oldMonthDF$monthName, levels = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec"))
# 
# midMonthDF$monthName = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec")
# midMonthDF$monthName = factor(midMonthDF$monthName, levels = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec"))

latestMonthDF$monthName = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec")
latestMonthDF$monthName = factor(latestMonthDF$monthName, levels = c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov","Dec"))



# # Make the three panel bar plot showing the anomaly
# oldPlot = ggplot(data = oldMonthDF) +
#   geom_col(mapping = aes(x=monthName, y=TempAnom), fill="light blue") +
#   xlab("1900 - 1909 Compared to 1981 - 2010") +
#   ylab("Anomaly (deg F)")
# 
# midPlot = ggplot(data = midMonthDF) +
#   geom_col(mapping = aes(x=monthName, y=TempAnom), fill="blue") +
#   xlab("1960 - 1969 Compared to 1981 - 2010") +
#   ylab("Anomaly (deg F)")
# 
# latestPlot = ggplot(data = latestMonthDF) +
#   geom_col(mapping = aes(x=monthName, y=TempAnom), fill="dark blue") +
#   xlab("2010 - 2019 Compared to 1981 - 2010") +
#   ylab("Anomaly (deg F)")
# 
# grid.arrange(oldPlot,midPlot,latestPlot, ncol = 1)

ggplot(data = latestMonthDF) +
  geom_col(mapping = aes(x=monthName, y=TempAnom), fill="dark blue") +
  xlab("2010 - 2019 Monthly Average Temperature Compared to the Whole Record") +
  ylab("Anomaly (deg F)")


```


